---
title: "Marker bat acoustic models draft 2"
output: html_document
date: "2022-11-28"
---

```{r}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(tidy = "styler")
options(knitr.table.format = "html")

# set project working directory according to the user system info 
# Otherwise, all the data (inputs and outputs, including figures) can be stored on a shared OneDrive folder

user <- Sys.info()['effective_user'] 
user
# this should print your nmbu user name - "apmc" in my case. 

wd <- getwd()
wd
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/Documents/1. PhD_Main/GitHub_link/MarkerBatAcousticAnalyses/MarkerBatAcousticAnalyses2022"



### Here is some example code from Pierre that I would like to implement soon which allows us to adjust all the directories based on the user

# if(Sys.info()['user'] == 'pidu') {                                                                                                                                               ## Pierre
#   gitDir <- 'C:/myDocuments/AlpineWolf'
#   dataDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/01_Data'  
#   analysisDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/02_Analysis'
#   simulationDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/03_Simulations'
#   meetingDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/04_Meetings'
#   reportDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/06_Report'
# } else if(Sys.info()['user'] == 'virginia') {                                                                                                                                ## Virginia
#   gitDir <- '/Users/virginia/Dropbox/Mac/Documents/GitHub/AlpineWolf'
#   dataDir <- '/Users/virginia/Dropbox/AlpineWolf/01_Data'
#   analysisDir <- '/Users/virginia/Dropbox/AlpineWolf/02_Analysis'
#   simulationDir <- '/Users/virginia/Dropbox/AlpineWolf/03_Simulations'
#   meetingDir <- '/Users/virginia/Dropbox/AlpineWolf/04_Meetings'
#   reportDir <- '/Users/virginia/Dropbox/AlpineWolf/06_Report'
# 

##########################################################
#### Work environment set up ####
##########################################################
library(knitr)
library(data.table)
library(tidyverse)
library(beepr)
library(lubridate)
library(purrr)
#renv::install("rstudio/renv")
library(renv)
library(stringr)
library(janitor)
library(anytime)
library(kableExtra)
library(papeR)
library(skimr)
library(vtable)
library(gratia)
library(DHARMa)
library(mgcv)
library(tidymv)

##########################################################
#### Import data, set up directories   ####
##########################################################

# All input data can be found on a shared OneDrive folder - we can both share the same input folder but we should have different Output folders. 

# for Katrine 
#input <- ""

# for Reed 
input <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Input/forModels"

dataset1 <- "guild_behavior_summarytable_site.csv" 
# df2C for the dataset aggregated by guild and bahavior from Marker aggregated to night_all bats
dataset2 <- "totalbatpass_summarytable_withbinary_batpass_night_aggregated_site data.csv"
# df2C for the dataset aggregated to all batsa from Marker aggregated to night_all bats
dataset3 <- "nightlyaggregated_zeros_binary_behavior and guild.csv"
# dataset aggregated to guild and behavior before being table transformed 
dataset4 <- "nightlyaggregatedMarker2020_totalbats_zeroinserted_weather_binary.csv"
# dataset aggregated to night by total bats before table transformed 
dataset5 <- "guild_behavior_batpass_summarytable_trimmed MRE social and met tower.csv"


path1 <- str_c(input, "/", dataset1)
path2 <- str_c(input, "/", dataset2)
path3 <- str_c(input, "/", dataset3)
path4 <- str_c(input, "/", dataset4)
path5 <- str_c(input, "/", dataset5)

bats_gb <- read_csv(path1) # 11412 obs of 20 variables 
bats_tot <- read.csv(path2) # # 951 obs of 18 variables variables
bats_gb_simple <- read.csv(path3) # 11412 obs of 17 vars 
bats_tot_simple <- read.csv(path4) # 951 obs of 15 variables
bats_gb_trim <- read.csv(path5) # 4866 obs of 20 variables 

# for Reed
output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Reed/Outputs"

# for Katrine 
#output <- ""

## 
 file.name <- "Step3.Marker bat acoustics models draft 2"
# 
 todays_date <- Sys.Date()
# 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
 dir.name
# 
# dir.create(dir.name) # be careful not to recreate existing directories 

output_today <- dir.name
output_today

```

# Recommended Youtube videos on GAM to understand the R code below

 https://www.youtube.com/watch?v=q4_t8jXcQgc
 https://www.youtube.com/watch?v=sgw4cu8hrZM&t=4038s
 
 #############################################################################
 ################ "bats_tot" ################
 #############################################################################
 
The "bats_tot" dataset 
Nightly aggregated total bat activity 

# Preparing the dataset for modeling 
Starting with bats total 
```{r}
str(bats_tot) #Seems like our factor variables are back to 'chr' variables again....
names(bats_tot)
#Housekeeping to make character variables factors
bats_tot$Habitat <- factor(bats_tot$Habitat)
bats_tot$Locality <- factor(bats_tot$Locality)
bats_tot$Site <- factor(bats_tot$Site)
bats_tot$Facility <- factor(bats_tot$Facility) 
bats_tot$jnight <- yday(bats_tot$night)

#Explore relationships and look check for outliers again
par(mfrow = c(3,2))
plot(bats_tot$jnight, bats_tot$batpass_prop)            
plot(bats_tot$temp_mean, bats_tot$batpass_prop)
plot(bats_tot$temp_max, bats_tot$batpass_prop)
plot(bats_tot$wind_mean, bats_tot$batpass_prop)
plot(bats_tot$wind_max, bats_tot$batpass_prop)

#The batspass_prop variable is "number of hours with >0 batpass" / "total number of observation hours" - per night
#This is your response, but often proportion variables cannot be modelled directly 
```
 
#Steps to prepare binary response (1=batpass, 0 = no batpass)
```{r}
# This is now only nightly aggregated so I am not sure this has the same intended effect... 

bats_tot$yes_batpass <- bats_tot$batpass01_sum   #batpass01_sum is sum of '1' values in batpass01, i.e. number of observ. hours per night when batpass recorded
bats_tot$no_batpass <- bats_tot$batpass01_length - bats_tot$batpass01_sum #batpass01_length is number of observ. hours per night

summary(bats_tot)

```

#Fitting he full model (I recommend not making the model more complex than this)
#Will have to do trial-and-error to get the k right


```{r}
m1 <- gam(cbind(yes_batpass,no_batpass) ~
              s(Locality, bs = "re") +                       #bs = "re" is equivalent to adding Locality as random intercept in a mixed model (GAMM)
              s(jnight, by = Habitat, bs = "gp", k =90) +    #allows for separate shapes of relationships with jnight for each habitat, bs = "gp" takes care of temporal autocorrelation
              Habitat +                                      #main effect of habitat (testing if means differ between habitats)
              s(temp_mean) +                                 #does not seem to make much difference whether I use max or mean temp, so I chose mean (feels less like cherry-picking, I think...)
              s(wind_mean) +                                 #same as for temp
              ti(temp_mean,wind_mean) ,                      #interaction between temp and wind
              data = bats_tot, method = "REML",                  #REML is not default, but is highly recommended by experts
              family = binomial, select=TRUE)                #may have to shift to quasibinomial if overdispersion, select = TRUE gives you automatic model selection

summary(m1) # not *quite* all terms significant - not Locality or jnight*Habitat 
# output below 
```
________________________________________________________________________________
summary(m1):
Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 90) + Habitat + 
    s(temp_mean) + s(wind_mean) + ti(temp_mean, wind_mean)

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)       -0.03617    0.22151  -0.163     0.87    
HabitatNatural     2.55800    0.29955   8.539  < 2e-16 ***
HabitatTurbinePad  3.24277    0.46442   6.982  2.9e-12 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                                           edf Ref.df Chi.sq  p-value    
s(Locality)                          0.3433791      6  0.358 0.386901    
s(jnight):HabitatMeteorologicalTower 0.0000476     72  0.000 0.877886    
s(jnight):HabitatNatural             2.5330933     87  8.624 0.012154 *  
s(jnight):HabitatTurbinePad          5.6185246     88 40.312  < 2e-16 ***
s(temp_mean)                         2.2539415      9 20.685 1.02e-06 ***
s(wind_mean)                         3.2134647      9 20.946 7.11e-06 ***
ti(temp_mean,wind_mean)              4.8139882     16 17.980 0.000105 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.369   Deviance explained = 37.5%
-REML = 319.49  Scale est. = 1         n = 951
________________________________________________________________________________


### Katrine: 
The reason why I use GAM with a rather simple random effect instead of GAMM, is that there are fewer options (e.g., distributions) available in GAMM

Model validation (may also use dHARMA to get more fancy validation plots)
Note that validation if binomial models is usually a pain and does not give meaningful validation plots
However, it works in our case, because of the aggregated nature of the data :-)
I found a similar example in https://www.youtube.com/watch?v=sgw4cu8hrZM&t=4038s 


```{r}
par(mfrow = c(2,2))
gam.check(m1)           #look at the plots, but also the output to check if k needs to be adjusted
gam.check(m1, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

overdispersion.m1 <- sum( residuals(m1, "pearson")^2 ) / m1$df.residual
overdispersion.m1
# 0.80653 - looking good! 

# From Katrine's Step3B GAM modelling script: 
#[1] 1.022701    #this value should ideally be 0.8-1.2, but it's not very bad!

```


# Basis dimension (k) checking results: 
The k-index should be approx. 1 or larger
Play play around with k....

```{r}
#plot(m1, pages=1,scheme=2,shade=TRUE)
bmp(file.path(output_today, "checking k dimensions m1 bats_tot.png"), width = 9, height = 6, units = "in", res = 350)

plot(m1, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m1, scales ="free")         #function 'draw' is from package 'gratia'
draw(m1, scales ="fixed")

dev.off() 
```

# m2 
# model temperature as linear effect

```{r}
m2 <- gam(cbind(yes_batpass,no_batpass) ~
            s(Locality, bs = "re") +                      
            s(jnight, by = Habitat, bs = "gp", k =90) +    
            Habitat +                                      
            temp_mean +                                 
            s(wind_mean) +                                 
            ti(temp_mean,wind_mean) ,                      
          data = bats_tot, method = "REML",                  
          family = binomial, select=TRUE)              
summary(m2) 

```
________________________________________________________________________________
summary(m2): 
Method: REML   Optimizer: outer newton
full convergence after 18 iterations.
Gradient range [-9.303033e-05,4.008943e-05]
(score 320.4746 & scale 1).
Hessian positive definite, eigenvalue range [2.88257e-06,0.7431721].
Model rank =  304 / 304 

Basis dimension (k) checking results. Low p-value (k-index<1) may
indicate that k is too low, especially if edf is close to k'.

                                           k'      edf k-index p-value    
s(Locality)                          8.00e+00 3.41e-01      NA      NA    
s(jnight):HabitatMeteorologicalTower 8.90e+01 6.45e-05    0.95    0.20    
s(jnight):HabitatNatural             8.90e+01 2.84e+00    0.95    0.20    
s(jnight):HabitatTurbinePad          8.90e+01 6.07e+00    0.95    0.22    
s(wind_mean)                         9.00e+00 3.24e+00    0.94    0.09 .  
ti(temp_mean,wind_mean)              1.60e+01 5.35e+00    0.85  <2e-16 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
________________________________________________________________________________

Only ti(temp_mean, wind_mean) is significant... and now Locality is NA, why would that be? 

```{r}

par(mfrow = c(2,2))
gam.check(m2, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

# Method: REML   Optimizer: outer newton
# full convergence after 18 iterations.
# Gradient range [-9.303033e-05,4.008943e-05]
# (score 320.4746 & scale 1).
# Hessian positive definite, eigenvalue range [2.88257e-06,0.7431721].
# Model rank =  304 / 304 
# 
# Basis dimension (k) checking results. Low p-value (k-index<1) may
# indicate that k is too low, especially if edf is close to k'.
# 
#                                            k'      edf k-index p-value    
# s(Locality)                          8.00e+00 3.41e-01      NA      NA    
# s(jnight):HabitatMeteorologicalTower 8.90e+01 6.45e-05    0.95    0.22    
# s(jnight):HabitatNatural             8.90e+01 2.84e+00    0.95    0.23    
# s(jnight):HabitatTurbinePad          8.90e+01 6.07e+00    0.95    0.24    
# s(wind_mean)                         9.00e+00 3.24e+00    0.94    0.09 .  
# ti(temp_mean,wind_mean)              1.60e+01 5.35e+00    0.85  <2e-16 ***
# ---
# Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

overdispersion.m2 <- sum( residuals(m2, "pearson")^2 ) / m2$df.residual
overdispersion.m2 
# Now at [1] 0.7804294, before was #[1] 1.02326 in Katrine's original script 


#Seemed like very high wind speeds creates spurious results
bats_prune<- bats_tot %>% filter(wind_mean < 12)
dim(bats)
dim(bats_prune)
# 937  21 # doqn from 951 obs (removed site nights with average wind speeds over 12 m/s)

#This pruning removed:
removed <- 100 - (100* (937/951))
removed #1.47% of data excluded (previously 1.7%)

```


# Re-fit with pruned dataset
```{r}

m2b <- gam(cbind(yes_batpass,no_batpass) ~
            s(Locality, bs = "re") +                      
            s(jnight, by = Habitat, bs = "gp", k =85) +    
            Habitat +                                      
            temp_mean +                                 
            s(wind_mean, k=12) +                                 
            ti(temp_mean,wind_mean) ,                      
          data = bats_prune, method = "REML",                  
          family = binomial, select=TRUE)              
summary(m2b) 

```
________________________________________________________________________________
Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 85) + Habitat + 
    temp_mean + s(wind_mean, k = 12) + ti(temp_mean, wind_mean)

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)       -4.42141    0.90313  -4.896 9.80e-07 ***
HabitatNatural     2.55878    0.35803   7.147 8.88e-13 ***
HabitatTurbinePad  3.21698    0.48662   6.611 3.82e-11 ***
temp_mean          0.29934    0.05491   5.451 5.00e-08 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                                           edf Ref.df Chi.sq p-value    
s(Locality)                          1.467e+00      6  1.864  0.2618    
s(jnight):HabitatMeteorologicalTower 5.539e-05     68  0.000  0.7122    
s(jnight):HabitatNatural             2.308e+00     82  6.090  0.0438 *  
s(jnight):HabitatTurbinePad          4.930e+00     83 37.827  <2e-16 ***
s(wind_mean)                         9.484e-05     11  0.000  0.6290    
ti(temp_mean,wind_mean)              8.058e+00     16 54.407  <2e-16 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.369   Deviance explained = 37.6%
-REML = 308.23  Scale est. = 1         n = 937
________________________________________________________________________________

The Locality factor is no longer NA

strong evidence of positive effect: 
s(jnight):HabitatTurbinePad
 ti(temp_mean,wind_mean)
 
 some evidence of positive effect:
 s(jnight):HabitatNatural 

```{r}
par(mfrow = c(2,2))
gam.check(m2b, rep=500) 

overdispersion.m2b <- sum( residuals(m2b, "pearson")^2 ) / m2b$df.residual
overdispersion.m2b
#Now 0.8011895, was 1.034198

#k-index is still a bit too low for all variables, but I move on for now to demonstrate prediction (sketching results)

bmp(file.path(output_today, "checking k dimensions m2b bats_tot.png"), width = 9, height = 6, units = "in", res = 350)
plot(m2b, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m2b, scales ="free")         #function 'draw' is from package 'gratia'
draw(m2b, scales ="fixed")

dev.off() 

# take a closer look at wind and temperature

bmp(file.path(output_today, "view gam for wind and temp m2b bats_tot.png"), width = 9, height = 6, units = "in", res = 350)

par(mfrow = c(1,2))
vis.gam(m2b, view = c("temp_mean", "wind_mean"))    #Not a very elegant plot, you can find prettier solutions, but helps understand the interaction between tempp and wind
vis.gam(m2b, view = c("wind_mean", "temp_mean"))

dev.off()


```

#Scetching the results - difference between Habitats
```{r}
levels(bats_prune$Habitat)
# "MeteorologicalTower" "Natural"             "TurbinePad"   

levels(bats_prune$Locality)
# [1] "MeteorologicalTower" "Turbine10"           "Turbine11"           "Turbine14"          
# [5] "Turbine2"            "Turbine4"            "Turbine8"            "Turbine9"

pdata <- with(bats_prune,
              expand.grid(temp_mean=14,
                          wind_mean = 6,
                          Habitat = c("Natural", "TurbinePad"),
                          Locality = c("Turbine14"),
                          jnight = seq(min(jnight), max(jnight), 1))) #1 to keep the integer format

head(pdata)
tail(pdata)

fit <- data.frame(predict(m2b, newdata=pdata, se.fit=TRUE, type = 'response'))
fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit))
pred <- cbind(pdata,fit)

head(pred)

#write.csv(pred, file.path(output_today, "prediction_m2b.csv"))

bmp(file.path(output_today, "predictions m2b bats_tot.png"), width = 9, height = 6, units = "in", res = 350)

#windows()
plt2 <- ggplot(pred, aes(x=jnight, y = fit, group = factor(Habitat))) +
    
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) +
    geom_line() + facet_wrap(Locality ~ Habitat, scales ='free_y') +
    labs(x = "Julian night", y ="prop of hours per night with batpass>0")
plt2

dev.off()

```

# Predicting wind at different levels of temperature
```{r}

pdata2 <- with(bats_prune,
              expand.grid(wind_mean = seq(min(wind_mean), max(wind_mean), length = 300),
                          temp_mean = c(12,15,20),
                          Habitat = c("TurbinePad"),
                          Locality = c("Turbine14"),
                          jnight = mean(jnight)))


head(pdata2)

fit <- data.frame(predict(m2b, newdata=pdata2, se.fit=TRUE, type = 'response'))
fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit))
pred2 <- cbind(pdata2,fit)

head(pred2)


new_labels <- c("12" = "temp 12", "15" = "temp 15", "20" = "temp 20")

bmp(file.path(output_today, "prediction of wind at dif temperatures m2b bats_tot.png"), width = 9, height = 6, units = "in", res = 350)

plt4 <- ggplot(pred2, aes(x=wind_mean, y = fit, group = factor(temp_mean))) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) +
  geom_line() + facet_wrap(~ temp_mean,scales ='fixed', labeller = labeller(temp_mean = new_labels)) +
  labs(x = "Average wind speed m/s", y ="Proportion of hours per night with batpass>0")
plt4

dev.off()
```


# Temp at different levels of wind
```{r}
pdata <- with(bats_prune,
              expand.grid(temp_mean = seq(min(temp_mean), max(temp_mean), length = 300),
                          wind_mean = c(4,7,10),
                          Habitat = c("TurbinePad"),
                          Locality = c("Turbine14"),
                          jnight = 212))


head(pdata)

fit <- data.frame(predict(m2b, newdata=pdata, se.fit=TRUE, type = 'response'))
fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit))
pred <- cbind(pdata,fit)

head(pred)

new_labels <- c("4" = "wind 4m/s", "7" = "wind 7m/s", "10" = "wind 10m/s")

bmp(file.path(output_today, "prediction of temp at dif wind speeds m2b bats_tot.png"), width = 9, height = 6, units = "in", res = 350) # test 

#windows()
plt4 <- ggplot(pred, aes(x=temp_mean, y = fit, group = factor(wind_mean))) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) +
  geom_line() + facet_wrap(~ wind_mean,scales ='fixed', labeller = labeller(wind_mean = new_labels)) +
  labs(x = expression(paste("Temperature [",degree,"C]")), y ="Proportion of hours per night with batpass>0")
plt4 #Ouch the degree synbol turned out as a square...need to fix that at some point

dev.off()
```
 
 From Katrine: 
<!-- #But it is really a hassel to have to predict for levels of Locality... -->
<!-- #One trick is to swap Loaclity with Facility in the mp model and refit the model -->
<!-- #Then you get away with predicting for Facility South (or North) -->
<!-- #But there is a way of excluding terms when predicting, see: https://cran.r-project.org/web/packages/tidymv/vignettes/predict-gam.html -->
<!-- #But I have not had time to figure this out yet. -->





