---
title: "Marker bat acoustic models draft 2"
output: html_document
date: "2022-11-28"
---

```{r}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(tidy = "styler")
options(knitr.table.format = "html")

# set project working directory according to the user system info 
# Otherwise, all the data (inputs and outputs, including figures) can be stored on a shared OneDrive folder

user <- Sys.info()['effective_user'] 
user
# this should print your nmbu user name - "apmc" in my case. 

wd <- getwd()
wd
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/Documents/1. PhD_Main/GitHub_link/MarkerBatAcousticAnalyses/MarkerBatAcousticAnalyses2022"



### Here is some example code from Pierre that I would like to implement soon which allows us to adjust all the directories based on the user

# if(Sys.info()['user'] == 'pidu') {                                                                                                                                               ## Pierre
#   gitDir <- 'C:/myDocuments/AlpineWolf'
#   dataDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/01_Data'  
#   analysisDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/02_Analysis'
#   simulationDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/03_Simulations'
#   meetingDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/04_Meetings'
#   reportDir <- 'C:/Users/pidu/Dropbox (AQEG)/AQEG Team Folder/AlpineWolf/06_Report'
# } else if(Sys.info()['user'] == 'virginia') {                                                                                                                                ## Virginia
#   gitDir <- '/Users/virginia/Dropbox/Mac/Documents/GitHub/AlpineWolf'
#   dataDir <- '/Users/virginia/Dropbox/AlpineWolf/01_Data'
#   analysisDir <- '/Users/virginia/Dropbox/AlpineWolf/02_Analysis'
#   simulationDir <- '/Users/virginia/Dropbox/AlpineWolf/03_Simulations'
#   meetingDir <- '/Users/virginia/Dropbox/AlpineWolf/04_Meetings'
#   reportDir <- '/Users/virginia/Dropbox/AlpineWolf/06_Report'
# 

##########################################################
#### Work environment set up ####
##########################################################
library(knitr)
library(data.table)
library(tidyverse)
library(beepr)
library(lubridate)
library(purrr)
#renv::install("rstudio/renv")
library(renv)
library(stringr)
library(janitor)
library(anytime)
library(kableExtra)
library(papeR)
library(skimr)
library(vtable)
library(gratia)
library(DHARMa)
library(mgcv)
library(tidymv)

##########################################################
#### Import data, set up directories   ####
##########################################################

# All input data can be found on a shared OneDrive folder - we can both share the same input folder but we should have different Output folders. 

# for Katrine 
#input <- ""

# for Reed 
input <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Input/forModels"

dataset1 <- "guild_behavior_summarytable_site.csv" 
# df2C for the dataset aggregated by guild and bahavior from Marker aggregated to night_all bats
dataset2 <- "totalbatpass_summarytable_withbinary_batpass_night_aggregated_site data.csv"
# df2C for the dataset aggregated to all batsa from Marker aggregated to night_all bats
dataset3 <- "nightlyaggregated_zeros_binary_behavior and guild.csv"
# dataset aggregated to guild and behavior before being table transformed 
dataset4 <- "nightlyaggregatedMarker2020_totalbats_zeroinserted_weather_binary.csv"
# dataset aggregated to night by total bats before table transformed 
dataset5 <- "guild_behavior_batpass_summarytable_trimmed MRE social and met tower.csv"


path1 <- str_c(input, "/", dataset1)
path2 <- str_c(input, "/", dataset2)
path3 <- str_c(input, "/", dataset3)
path4 <- str_c(input, "/", dataset4)
path5 <- str_c(input, "/", dataset5)

bats_gb <- read_csv(path1) # 11412 obs of 20 variables 
bats_tot <- read.csv(path2) # # 951 obs of 18 variables variables
bats_gb_simple <- read.csv(path3) # 11412 obs of 17 vars 
bats_tot_simple <- read.csv(path4) # 951 obs of 15 variables
bats_gb_trim <- read.csv(path5) # 4866 obs of 20 variables 

# for Reed
output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Reed/Outputs"

# for Katrine 
#output <- ""

## 
 file.name <- "Step3.Marker bat acoustics models draft 2"
# 
 todays_date <- Sys.Date()
# 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
 dir.name
# 
 dir.create(dir.name) # be careful not to recreate existing directories 

output_today <- dir.name
output_today

```

# Recommended Youtube videos on GAM to understand the R code below

 https://www.youtube.com/watch?v=q4_t8jXcQgc
 https://www.youtube.com/watch?v=sgw4cu8hrZM&t=4038s
 
 #############################################################################
 ################ "bats_tot" ################
 #############################################################################
 
The "bats_tot" dataset 
Nightly aggregated total bat activity 

# Preparing the dataset for modeling 
Starting with bats total 
```{r}
str(bats_tot) #Seems like our factor variables are back to 'chr' variables again....
names(bats_tot)
#Housekeeping to make character variables factors
bats_tot$Habitat <- factor(bats_tot$Habitat)
bats_tot$Locality <- factor(bats_tot$Locality)
bats_tot$Site <- factor(bats_tot$Site)
bats_tot$Facility <- factor(bats_tot$Facility) 
bats_tot$jnight <- yday(bats_tot$night)

#Explore relationships and look check for outliers again
par(mfrow = c(3,2))
plot(bats_tot$jnight, bats_tot$batpass_prop)            
plot(bats_tot$temp_mean, bats_tot$batpass_prop)
plot(bats_tot$temp_max, bats_tot$batpass_prop)
plot(bats_tot$wind_mean, bats_tot$batpass_prop)
plot(bats_tot$wind_max, bats_tot$batpass_prop)

#The batspass_prop variable is "number of hours with >0 batpass" / "total number of observation hours" - per night
#This is your response, but often proportion variables cannot be modelled directly 
```
 
#Steps to prepare binary response (1=batpass, 0 = no batpass)
```{r}
# This is now only nightly aggregated so I am not sure this has the same intended effect... 

bats_tot$yes_batpass <- bats_tot$batpass01_sum   #batpass01_sum is sum of '1' values in batpass01, i.e. number of observ. hours per night when batpass recorded
bats_tot$no_batpass <- bats_tot$batpass01_length - bats_tot$batpass01_sum #batpass01_length is number of observ. hours per night

summary(bats_tot)

```

#Fitting he full model (I recommend not making the model more complex than this)
#Will have to do trial-and-error to get the k right

```{r}
m1 <- gam(cbind(yes_batpass,no_batpass) ~
              s(Locality, bs = "re") +                       #bs = "re" is equivalent to adding Locality as random intercept in a mixed model (GAMM)
              s(jnight, by = Habitat, bs = "gp", k =90) +    #allows for separate shapes of relationships with jnight for each habitat, bs = "gp" takes care of temporal autocorrelation
              Habitat +                                      #main effect of habitat (testing if means differ between habitats)
              s(temp_mean) +                                 #does not seem to make much difference whether I use max or mean temp, so I chose mean (feels less like cherry-picking, I think...)
              s(wind_mean) +                                 #same as for temp
              ti(temp_mean,wind_mean) ,                      #interaction between temp and wind
              data = bats_tot, method = "REML",                  #REML is not default, but is highly recommended by experts
              family = binomial, select=TRUE)                #may have to shift to quasibinomial if overdispersion, select = TRUE gives you automatic model selection

summary(m1) # not *quite* all terms significant - not Locality or jnight*Habitat 
# output below 
```
________________________________________________________________________________
summary(m1):
Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 90) + Habitat + 
    s(temp_mean) + s(wind_mean) + ti(temp_mean, wind_mean)

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)       -0.03617    0.22151  -0.163     0.87    
HabitatNatural     2.55800    0.29955   8.539  < 2e-16 ***
HabitatTurbinePad  3.24277    0.46442   6.982  2.9e-12 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                                           edf Ref.df Chi.sq  p-value    
s(Locality)                          0.3433791      6  0.358 0.386901    
s(jnight):HabitatMeteorologicalTower 0.0000476     72  0.000 0.877886    
s(jnight):HabitatNatural             2.5330933     87  8.624 0.012154 *  
s(jnight):HabitatTurbinePad          5.6185246     88 40.312  < 2e-16 ***
s(temp_mean)                         2.2539415      9 20.685 1.02e-06 ***
s(wind_mean)                         3.2134647      9 20.946 7.11e-06 ***
ti(temp_mean,wind_mean)              4.8139882     16 17.980 0.000105 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.369   Deviance explained = 37.5%
-REML = 319.49  Scale est. = 1         n = 951
________________________________________________________________________________


### Katrine: 
The reason why I use GAM with a rather simple random effect instead of GAMM, is that there are fewer options (e.g., distributions) available in GAMM

Model validation (may also use dHARMA to get more fancy validation plots)
Note that validation if binomial models is usually a pain and does not give meaningful validation plots
However, it works in our case, because of the aggregated nature of the data :-)
I found a similar example in https://www.youtube.com/watch?v=sgw4cu8hrZM&t=4038s 


```{r}
par(mfrow = c(2,2))
gam.check(m1)           #look at the plots, but also the output to check if k needs to be adjusted
gam.check(m1, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

overdispersion.m1 <- sum( residuals(m1, "pearson")^2 ) / m1$df.residual
overdispersion.m1
# 0.80653 - looking good! 

# From Katrine's Step3B GAM modelling script: 
#[1] 1.022701    #this value should ideally be 0.8-1.2, but it's not very bad!

```


# Basis dimension (k) checking results: 
The k-index should be approx. 1 or larger
Play play around with k....

```{r}
#plot(m1, pages=1,scheme=2,shade=TRUE)
bmp(file.path(output_today, "checking k dimensions m1 bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

plot(m1, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m1, scales ="free")         #function 'draw' is from package 'gratia'
draw(m1, scales ="fixed")

dev.off() 
```

# m2 
# model temperature as linear effect

```{r}
m2 <- gam(cbind(yes_batpass,no_batpass) ~
            s(Locality, bs = "re") +                      
            s(jnight, by = Habitat, bs = "gp", k =90) +    
            Habitat +                                      
            temp_mean +                                 
            s(wind_mean) +                                 
            ti(temp_mean,wind_mean) ,                      
          data = bats_tot, method = "REML",                  
          family = binomial, select=TRUE)              
summary(m2) 

```
________________________________________________________________________________
summary(m2): 
Method: REML   Optimizer: outer newton
full convergence after 18 iterations.
Gradient range [-9.303033e-05,4.008943e-05]
(score 320.4746 & scale 1).
Hessian positive definite, eigenvalue range [2.88257e-06,0.7431721].
Model rank =  304 / 304 

Basis dimension (k) checking results. Low p-value (k-index<1) may
indicate that k is too low, especially if edf is close to k'.

                                           k'      edf k-index p-value    
s(Locality)                          8.00e+00 3.41e-01      NA      NA    
s(jnight):HabitatMeteorologicalTower 8.90e+01 6.45e-05    0.95    0.20    
s(jnight):HabitatNatural             8.90e+01 2.84e+00    0.95    0.20    
s(jnight):HabitatTurbinePad          8.90e+01 6.07e+00    0.95    0.22    
s(wind_mean)                         9.00e+00 3.24e+00    0.94    0.09 .  
ti(temp_mean,wind_mean)              1.60e+01 5.35e+00    0.85  <2e-16 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
________________________________________________________________________________

Only ti(temp_mean, wind_mean) is significant... and now Locality is NA, why would that be? 

```{r}

par(mfrow = c(2,2))
gam.check(m2)
gam.check(m2, rep=500)  
#rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

# Method: REML   Optimizer: outer newton
# full convergence after 18 iterations.
# Gradient range [-9.303033e-05,4.008943e-05]
# (score 320.4746 & scale 1).
# Hessian positive definite, eigenvalue range [2.88257e-06,0.7431721].
# Model rank =  304 / 304 
# 
# Basis dimension (k) checking results. Low p-value (k-index<1) may
# indicate that k is too low, especially if edf is close to k'.
# 
#                                            k'      edf k-index p-value    
# s(Locality)                          8.00e+00 3.41e-01      NA      NA    
# s(jnight):HabitatMeteorologicalTower 8.90e+01 6.45e-05    0.95    0.22    
# s(jnight):HabitatNatural             8.90e+01 2.84e+00    0.95    0.23    
# s(jnight):HabitatTurbinePad          8.90e+01 6.07e+00    0.95    0.24    
# s(wind_mean)                         9.00e+00 3.24e+00    0.94    0.09 .  
# ti(temp_mean,wind_mean)              1.60e+01 5.35e+00    0.85  <2e-16 ***
# ---
# Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1


### Interpretting gam.checl() output
## First check if the model converges - this does after 21 iterations!
## If it didn't then the results would not be meaningful 
## Now each smooth term is evaluated to see if the residuals of these terms are randomly distributed. 
## If the p value is low for a smooth term, then the residuals are not random and most likely the K value should be increased. 
## In this case, the p value of tensor product interaction term (ti) for  temp and mean is very low and the current k value is 16. Notice that the other k values on the smooth terms are either 9 or close to 90. 
## Tensor product smooths and interaction, playing with k for these types of terms in not the same as for normal smooth terms... need to specify both the kvalues i.e. k = c(4,5) 

## quick rundown on some of the lexicon behind tensor terms: 
#   te() is like x*Z or x + z + x:z if you want to write it all out
#   ti() is like X:Z only
## 

overdispersion.m2 <- sum( residuals(m2, "pearson")^2 ) / m2$df.residual
overdispersion.m2 
# Now at [1] 0.7804294, before was #[1] 1.02326 in Katrine's original script 


#Seemed like very high wind speeds creates spurious results
bats_prune<- bats_tot %>% filter(wind_mean < 12)
dim(bats)
dim(bats_prune)
# 937  21 # doqn from 951 obs (removed site nights with average wind speeds over 12 m/s)

#This pruning removed:
removed <- 100 - (100* (937/951))
removed #1.47% of data excluded (previously 1.7%)

```




<!-- # Re-fit with pruned dataset -->
<!-- ```{r} -->

<!-- m2b <- gam(cbind(yes_batpass,no_batpass) ~ -->
<!--             s(Locality, bs = "re") +                       -->
<!--             s(jnight, by = Habitat, bs = "gp", k =85) +     -->
<!--             Habitat +                                       -->
<!--             temp_mean +                                  -->
<!--             s(wind_mean, k=12) +                                  -->
<!--             ti(temp_mean,wind_mean) ,                       -->
<!--           data = bats_prune, method = "REML",                   -->
<!--           family = binomial, select=TRUE)               -->
<!-- summary(m2b)  -->

<!-- ``` -->
<!-- ________________________________________________________________________________ -->
<!-- Family: binomial  -->
<!-- Link function: logit  -->

<!-- Formula: -->
<!-- cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") +  -->
<!--     s(jnight, by = Habitat, bs = "gp", k = 85) + Habitat +  -->
<!--     temp_mean + s(wind_mean, k = 12) + ti(temp_mean, wind_mean) -->

<!-- Parametric coefficients: -->
<!--                   Estimate Std. Error z value Pr(>|z|)     -->
<!-- (Intercept)       -4.42141    0.90313  -4.896 9.80e-07 *** -->
<!-- HabitatNatural     2.55878    0.35803   7.147 8.88e-13 *** -->
<!-- HabitatTurbinePad  3.21698    0.48662   6.611 3.82e-11 *** -->
<!-- temp_mean          0.29934    0.05491   5.451 5.00e-08 *** -->
<!-- --- -->
<!-- Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 -->

<!-- Approximate significance of smooth terms: -->
<!--                                            edf Ref.df Chi.sq p-value     -->
<!-- s(Locality)                          1.467e+00      6  1.864  0.2618     -->
<!-- s(jnight):HabitatMeteorologicalTower 5.539e-05     68  0.000  0.7122     -->
<!-- s(jnight):HabitatNatural             2.308e+00     82  6.090  0.0438 *   -->
<!-- s(jnight):HabitatTurbinePad          4.930e+00     83 37.827  <2e-16 *** -->
<!-- s(wind_mean)                         9.484e-05     11  0.000  0.6290     -->
<!-- ti(temp_mean,wind_mean)              8.058e+00     16 54.407  <2e-16 *** -->
<!-- --- -->
<!-- Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 -->

<!-- R-sq.(adj) =  0.369   Deviance explained = 37.6% -->
<!-- -REML = 308.23  Scale est. = 1         n = 937 -->
<!-- ________________________________________________________________________________ -->

<!-- The Locality factor is no longer NA -->

<!-- strong evidence of positive effect:  -->
<!-- s(jnight):HabitatTurbinePad -->
<!--  ti(temp_mean,wind_mean) -->

<!--  some evidence of positive effect: -->
<!--  s(jnight):HabitatNatural  -->

<!-- ```{r} -->
<!-- par(mfrow = c(2,2)) -->
<!-- gam.check(m2b, rep=500)  -->

<!-- overdispersion.m2b <- sum( residuals(m2b, "pearson")^2 ) / m2b$df.residual -->
<!-- overdispersion.m2b -->
<!-- #Now 0.8011895, was 1.034198 -->

<!-- #k-index is still a bit too low for all variables, but I move on for now to demonstrate prediction (sketching results) -->

<!-- bmp(file.path(output_today, "checking k dimensions m2b bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350) -->
<!-- plot(m2b, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE) -->

<!-- draw(m2b, scales ="free")         #function 'draw' is from package 'gratia' -->
<!-- draw(m2b, scales ="fixed") -->

<!-- dev.off()  -->

<!-- # take a closer look at wind and temperature -->

<!-- bmp(file.path(output_today, "view gam for wind and temp m2b bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350) -->

<!-- par(mfrow = c(1,2)) -->
<!-- vis.gam(m2b, view = c("temp_mean", "wind_mean"))    #Not a very elegant plot, you can find prettier solutions, but helps understand the interaction between tempp and wind -->
<!-- vis.gam(m2b, view = c("wind_mean", "temp_mean")) -->

<!-- dev.off() -->


<!-- ``` -->

<!-- #Sketching the results - difference between Habitats -->
<!-- ```{r} -->
<!-- levels(bats_prune$Habitat) -->
<!-- # "MeteorologicalTower" "Natural"             "TurbinePad"    -->

<!-- levels(bats_prune$Locality) -->
<!-- # [1] "MeteorologicalTower" "Turbine10"           "Turbine11"           "Turbine14"           -->
<!-- # [5] "Turbine2"            "Turbine4"            "Turbine8"            "Turbine9" -->

<!-- pdata <- with(bats_prune, -->
<!--               expand.grid(temp_mean=14, -->
<!--                           wind_mean = 6, -->
<!--                           Habitat = c("Natural", "TurbinePad"), -->
<!--                           Locality = c("Turbine14"), -->
<!--                           jnight = seq(min(jnight), max(jnight), 1))) #1 to keep the integer format -->

<!-- head(pdata) -->
<!-- tail(pdata) -->

<!-- fit <- data.frame(predict(m2b, newdata=pdata, se.fit=TRUE, type = 'response')) -->
<!-- fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit)) -->
<!-- pred <- cbind(pdata,fit) -->

<!-- head(pred) -->

<!-- #write.csv(pred, file.path(output_today, "prediction_m2b.csv")) -->

<!-- bmp(file.path(output_today, "predictions m2b bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350) -->

<!-- ##windows() -->
<!-- plt2 <- ggplot(pred, aes(x=jnight, y = fit, group = factor(Habitat))) + -->

<!--     geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) + -->
<!--     geom_line() + facet_wrap(Locality ~ Habitat, scales ='free_y') + -->
<!--     labs(x = "Julian night", y ="prop of hours per night with batpass>0") -->
<!-- plt2 -->

<!-- dev.off() -->

<!-- ``` -->

<!-- # Predicting wind at different levels of temperature -->
<!-- ```{r} -->

<!-- pdata2 <- with(bats_prune, -->
<!--               expand.grid(wind_mean = seq(min(wind_mean), max(wind_mean), length = 300), -->
<!--                           temp_mean = c(12,15,20), -->
<!--                           Habitat = c("TurbinePad"), -->
<!--                           Locality = c("Turbine14"), -->
<!--                           jnight = mean(jnight))) -->


<!-- head(pdata2) -->

<!-- fit <- data.frame(predict(m2b, newdata=pdata2, se.fit=TRUE, type = 'response')) -->
<!-- fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit)) -->
<!-- pred2 <- cbind(pdata2,fit) -->

<!-- head(pred2) -->


<!-- new_labels <- c("12" = "temp 12", "15" = "temp 15", "20" = "temp 20") -->

<!-- bmp(file.path(output_today, "prediction of wind at dif temperatures m2b bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350) -->

<!-- plt4 <- ggplot(pred2, aes(x=wind_mean, y = fit, group = factor(temp_mean))) + -->
<!--   geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) + -->
<!--   geom_line() + facet_wrap(~ temp_mean,scales ='fixed', labeller = labeller(temp_mean = new_labels)) + -->
<!--   labs(x = "Average wind speed m/s", y ="Proportion of hours per night with batpass>0") -->
<!-- plt4 -->

<!-- dev.off() -->
<!-- ``` -->


<!-- # Temp at different levels of wind -->
<!-- ```{r} -->
<!-- pdata <- with(bats_prune, -->
<!--               expand.grid(temp_mean = seq(min(temp_mean), max(temp_mean), length = 300), -->
<!--                           wind_mean = c(4,7,10), -->
<!--                           Habitat = c("TurbinePad"), -->
<!--                           Locality = c("Turbine14"), -->
<!--                           jnight = 212)) -->


<!-- head(pdata) -->

<!-- fit <- data.frame(predict(m2b, newdata=pdata, se.fit=TRUE, type = 'response')) -->
<!-- fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit)) -->
<!-- pred <- cbind(pdata,fit) -->

<!-- head(pred) -->

<!-- new_labels <- c("4" = "wind 4m/s", "7" = "wind 7m/s", "10" = "wind 10m/s") -->

<!-- bmp(file.path(output_today, "prediction of temp at dif wind speeds m2b bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350) -->

<!-- ##windows() -->
<!-- plt4 <- ggplot(pred, aes(x=temp_mean, y = fit, group = factor(wind_mean))) + -->
<!--   geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) + -->
<!--   geom_line() + facet_wrap(~ wind_mean,scales ='fixed', labeller = labeller(wind_mean = new_labels)) + -->
<!--   labs(x = expression(paste("Temperature [",degree,"C]")), y ="Proportion of hours per night with batpass>0") -->
<!-- plt4 #Ouch the degree synbol turned out as a square...need to fix that at some point -->

<!-- dev.off() -->
<!-- ``` -->
 
 From Katrine: 
<!-- #But it is really a hassel to have to predict for levels of Locality... -->
<!-- #One trick is to swap Loaclity with Facility in the mp model and refit the model -->
<!-- #Then you get away with predicting for Facility South (or North) -->
<!-- #But there is a way of excluding terms when predicting, see: https://cran.r-project.org/web/packages/tidymv/vignettes/predict-gam.html -->
<!-- #But I have not had time to figure this out yet. -->

--------------------------------------------------------------------------------

# Experitmenting with different k levels for the total bat dataset 
m2 - for wind temp interaction, K = 90

# m2: y = total binary nightly bat pass, k = 90 for wind and temp 

```{r}
m2 <- gam(cbind(yes_batpass,no_batpass) ~
              s(Locality, bs = "re") +                      
              s(jnight, by = Habitat, bs = "gp", k =90) +    
              Habitat +                                      
              s(temp_mean) +                                 
              s(wind_mean) +                                 
              ti(temp_mean,wind_mean) ,                      
              data = bats_tot, method = "REML",                  
              family = binomial, select=TRUE)                #

summary(m2) 
```
________________________________________________________________________________

Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 80) + Habitat + 
    s(temp_mean) + s(wind_mean) + ti(temp_mean, wind_mean)

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)       -0.03617    0.22151  -0.163     0.87    
HabitatNatural     2.55799    0.29955   8.539  < 2e-16 ***
HabitatTurbinePad  3.24270    0.46439   6.983 2.89e-12 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                                           edf Ref.df Chi.sq  p-value    
s(Locality)                          3.434e-01      6  0.358 0.386902    
s(jnight):HabitatMeteorologicalTower 7.701e-05     65  0.000 0.877862    
s(jnight):HabitatNatural             2.533e+00     78  8.624 0.012154 *  
s(jnight):HabitatTurbinePad          5.617e+00     78 40.311  < 2e-16 ***
s(temp_mean)                         2.254e+00      9 20.685 1.02e-06 ***
s(wind_mean)                         3.213e+00      9 20.945 7.12e-06 ***
ti(temp_mean,wind_mean)              4.814e+00     16 17.980 0.000105 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.369   Deviance explained = 37.5%
-REML = 319.49  Scale est. = 1         n = 951

________________________________________________________________________________


```{r}
par(mfrow = c(2,2))
gam.check(m1)           #look at the plots, but also the output to check if k needs to be adjusted
gam.check(m1, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

overdispersion.m1 <- sum( residuals(m1, "pearson")^2 ) / m1$df.residual
overdispersion.m1
# 0.80653 - looking good! 

# From Katrine's Step3B GAM modelling script: 
#[1] 1.022701    #this value should ideally be 0.8-1.2, but it's not very bad!

```


# Basis dimension (k) checking results: 
The k-index should be approx. 1 or larger
Play play around with k....

```{r}
#plot(m1, pages=1,scheme=2,shade=TRUE)
bmp(file.path(output_today, "checking k dimensions m1 bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

plot(m1, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m1, scales ="free")         #function 'draw' is from package 'gratia'
draw(m1, scales ="fixed")

dev.off() 
```

# m2 
# model temperature as linear effect, assign higher k values to the ti() term 

```{r}
m2 <- gam(cbind(yes_batpass,no_batpass) ~
            s(Locality, bs = "re") +                      
            s(jnight, by = Habitat, bs = "gp", k =90) +    
            Habitat +                                      
            temp_mean +                                 
            s(wind_mean) +                                 
            ti(temp_mean,wind_mean, k = c(20, 20)) ,                      
          data = bats_tot, method = "REML",                  
          family = binomial, select=TRUE)              
summary(m2) 

```

summary(*m2*)
Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 90) + Habitat + 
    temp_mean + s(wind_mean) + ti(temp_mean, wind_mean, k = c(20, 
    20))

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)       -2.64438    1.16274  -2.274   0.0229 *  
HabitatNatural     2.71720    0.38548   7.049 1.80e-12 ***
HabitatTurbinePad  3.29057    0.48947   6.723 1.78e-11 ***
temp_mean          0.17623    0.07177   2.456   0.0141 *  
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                                           edf Ref.df Chi.sq  p-value    
s(Locality)                          1.620e+00      6  2.054  0.26070    
s(jnight):HabitatMeteorologicalTower 1.673e-04     72  0.000  0.59525    
s(jnight):HabitatNatural             2.627e+00     87  8.375  0.00845 ** 
s(jnight):HabitatTurbinePad          4.377e+00     88 32.974  < 2e-16 ***
s(wind_mean)                         7.670e+00      9 39.700  < 2e-16 ***
ti(temp_mean,wind_mean)              2.578e+01     81 57.695 4.35e-06 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.444   Deviance explained = 46.5%
-REML = 315.42  Scale est. = 1         n = 951

```{r}
windows()
par(mfrow = c(2,2))
gam.check(m2)           #look at the plots, but also the output to check if k needs to be adjusted
gam.check(m2, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

# Now the smooth terms pass the basis dimension test! 

overdispersion.m2 <- sum( residuals(m2, "pearson")^2 ) / m2$df.residual
overdispersion.m2
# 0.5493729- under dispersed:( 

# From Katrine's Step3B GAM modelling script: 
#[1] 1.022701    #this value should ideally be 0.8-1.2, but it's not very bad!

```


# Basis dimension (k) checking results: 
The k-index should be approx. 1 or larger
Play play around with k....

```{r}
#plot(m1, pages=1,scheme=2,shade=TRUE)
bmp(file.path(output_today, "checking k dimensions m2 bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)
windows()
plot(m2, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m2, scales ="free")         #function 'draw' is from package 'gratia'
draw(m2, scales ="fixed")
# wind is too wiggly, but there is no relationship for the met tower(probably too little data...) 
dev.off() 
```

# m3 
### Try once more with slightly lower k values for the ti() term 

```{r}
m3 <- gam(cbind(yes_batpass,no_batpass) ~
            s(Locality, bs = "re") +                      
            s(jnight, by = Habitat, bs = "gp", k =90) +    
            Habitat +                                      
            temp_mean +                                 
            s(wind_mean) +                                 
            ti(temp_mean,wind_mean, k = c(10, 10)) ,                      
          data = bats_tot, method = "REML",                  
          family = binomial, select=TRUE)              
summary(m3) 

```

summary(*m3*)
Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 90) + Habitat + 
    temp_mean + s(wind_mean) + ti(temp_mean, wind_mean, k = c(10, 
    10))

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)       -4.67455    0.94713  -4.935 8.00e-07 ***
HabitatNatural     2.58345    0.30044   8.599  < 2e-16 ***
HabitatTurbinePad  3.22108    0.44943   7.167 7.67e-13 ***
temp_mean          0.31458    0.05915   5.319 1.05e-07 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                                           edf Ref.df Chi.sq p-value    
s(Locality)                          2.415e-01      6  0.249  0.3928    
s(jnight):HabitatMeteorologicalTower 3.293e-04     69  0.000  0.6015    
s(jnight):HabitatNatural             2.762e+00     87  9.344  0.0089 ** 
s(jnight):HabitatTurbinePad          4.927e+00     88 38.850  <2e-16 ***
s(wind_mean)                         7.987e-05      9  0.000  0.2455    
ti(temp_mean,wind_mean)              1.751e+01     80 72.322  <2e-16 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.397   Deviance explained = 40.2%
-REML = 317.61  Scale est. = 1         n = 951

```{r}
windows()
par(mfrow = c(2,2))
gam.check(m3)           #look at the plots, but also the output to check if k needs to be adjusted
gam.check(m3, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

# Now the smooth terms pass the basis dimension test! 

overdispersion.m3 <- sum( residuals(m3, "pearson")^2 ) / m3$df.residual
overdispersion.m3
# 0.5493729- under dispersed:( 

# From Katrine's Step3B GAM modelling script: 
#[1] 1.022701    #this value should ideally be 0.8-1.2, but it's not very bad!

```


# Basis dimension (k) checking results: 
The k-index should be approx. 1 or larger
Play play around with k....

```{r}
#plot(m1, pages=1,scheme=2,shade=TRUE)
bmp(file.path(output_today, "checking k dimensions m3 bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)
windows()
plot(m3, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m3, scales ="free")         #function 'draw' is from package 'gratia'
draw(m3, scales ="fixed")
# wind is too wiggly, but there is no relationship for the met tower(probably too little data...) 
dev.off() 
```


# What did we learn?
It looks like model2, with the ti(temp_mean,wind_mean) k value adjusted to k = c(20,20) passes the gam.check test but then the model is a bit underdispersed. 

To Address under dispersion in binary GAMs:

https://dergipark.org.tr/en/download/article-file/922496 
This article suggests using an extended Altham distribution (EAD) family for under dispersed binomial count data- 
however I cannot find any commands in the mgcv package that will allow me to call that distribution.

The MM package does have the option to use this family but before changing packages, it is perhaps wiser to try and use counts as a response variable 

https://cran.r-project.org/web/packages/MM/vignettes/Gianfranco.pdf 




--------------------------------------------------------------------------------
Similar approach, now with 4 different data sets - 

LRE feeding
LRE commuting
SRE feeding
SRE commuting

Prepare new subsetted data
Take from bats_gb_trim (already removed met tower, MRE and social passes)'

```{r}

bats_gb_trim$Habitat <- factor(bats_gb_trim$Habitat)
bats_gb_trim$Locality <- factor(bats_gb_trim$Locality)
bats_gb_trim$Site <- factor(bats_gb_trim$Site)
bats_gb_trim$Facility <- factor(bats_gb_trim$Facility)
bats_gb_trim$guild <- factor(bats_gb_trim$guild)
bats_gb_trim$behavior <- factor(bats_gb_trim$behavior)
bats_gb_trim$jnight <- yday(bats_gb_trim$night)

bats_gb_trim$yes_batpass <- bats_gb_trim$batpass01_sum   #batpass01_sum is sum of '1' values in batpass01, i.e. number of observ. hours per night when batpass recorded
bats_gb_trim$no_batpass <- bats_gb_trim$batpass01_length - bats_gb_trim$batpass01_sum #batpass01_length is number of observ. hours per night

summary(bats_gb_trim)

LREf <- bats_gb_trim %>% filter(guild == "LRE", behavior == "Feeding") %>% droplevels()
LREc <- bats_gb_trim %>% filter(guild %in% "LRE", behavior %in% "Commuting") %>% droplevels()

SREf <- bats_gb_trim %>% filter(guild == "SRE", behavior == "Feeding") %>% droplevels()
SREc <- bats_gb_trim %>% filter(guild == "SRE", behavior == "Commuting") %>% droplevels()

summary(LREf) # 811 obs of 23 vars 
summary(LREc) # 811 obs of 23 vars 

summary(SREf) # 811 obs of 23 vars 
summary(SREc) # 811 obs of 23 vars 
```

# Now with LRE feeding

```{r}
m1_LREf <- gam(cbind(yes_batpass,no_batpass) ~
              s(Locality, bs = "re") +                       #bs = "re" is equivalent to adding Locality as random intercept in a mixed model (GAMM)
              s(jnight, by = Habitat, bs = "gp", k =90) +    
              Habitat +                                      
              s(temp_mean) +                               
              s(wind_mean) +                                 #same as for temp
              ti(temp_mean,wind_mean) ,                      #interaction between temp and wind
              data = LREf, method = "REML",                  #REML is not default, but is highly recommended by experts
              family = binomial, select=TRUE)                #may have to shift to quasibinomial if overdispersion, select = TRUE gives you automatic model selection

summary(m1_LREf) # 
# Now everything is significant besdies the ti(temp_mean, wind_mean) term
# Opposite from the last m1 model basically 
```
________________________________________________________________________________

Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 90) + Habitat + 
    s(temp_mean) + s(wind_mean) + ti(temp_mean, wind_mean)

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)        -1.2595     0.3560  -3.538 0.000403 ***
HabitatTurbinePad   0.6617     0.1992   3.321 0.000896 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                                  edf Ref.df Chi.sq  p-value    
s(Locality)                 5.5265823      6  55.65  < 2e-16 ***
s(jnight):HabitatNatural    1.8601064     87  18.16 6.89e-05 ***
s(jnight):HabitatTurbinePad 3.7275409     88  23.97 7.41e-05 ***
s(temp_mean)                3.2288598      9  72.42  < 2e-16 ***
s(wind_mean)                3.4219551      9  31.51  < 2e-16 ***
ti(temp_mean,wind_mean)     0.0004462     16   0.00    0.475    
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.372   Deviance explained = 33.6%
-REML = 393.62  Scale est. = 1         n = 811

________________________________________________________________________________

```{r}
par(mfrow = c(2,2))
gam.check(m1_LREf)           #look at the plots, but also the output to check if k needs to be adjusted
gam.check(m1_LREf, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

overdispersion.m1_LREf <- sum( residuals(m1_LREf, "pearson")^2 ) / m1_LREf$df.residual
overdispersion.m1_LREf
# 0.984761 ... it is alright! 

# From Katrine's Step3B GAM modelling script: 
#[1] 1.022701    #this value should ideally be 0.8-1.2, but it's not very bad!

```


# Basis dimension (k) checking results: 
The k-index should be approx. 1 or larger
Play play around with k....

```{r}

bmp(file.path(output_today, "checking k dimensions m1_LREf bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

#windows()
plot(m1_LREf, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m1_LREf, scales ="free")         #function 'draw' is from package 'gratia'
draw(m1_LREf, scales ="fixed")

dev.off() 
```


#Sketching the results - difference between Habitats
```{r}
levels(LREf$Habitat)
# "Natural"    "TurbinePad"

levels(LREf$Locality)
# [1] "Turbine10"           "Turbine11"           "Turbine14"          
# [5] "Turbine2"            "Turbine4"            "Turbine8"            "Turbine9"

pdata <- with(LREf,
              expand.grid(temp_mean=14,
                          wind_mean = 6,
                          Habitat = c("Natural", "TurbinePad"),
                          Locality = c("Turbine14"),
                          jnight = seq(min(jnight), max(jnight), 1))) #1 to keep the integer format

head(pdata)
tail(pdata)

fit <- data.frame(predict(m1_LREf, newdata=pdata, se.fit=TRUE, type = 'response'))
fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit))
pred <- cbind(pdata,fit)

head(pred)


bmp(file.path(output_today, "predictions m1_LREf bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

#windows()
plt2 <- ggplot(pred, aes(x=jnight, y = fit, group = factor(Habitat))) +
    
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) +
    geom_line() + facet_wrap(Locality ~ Habitat, scales ='free_y') +
    labs(x = "Julian night", y ="prop of hours per night with batpass>0")
plt2

dev.off()

```


# Now with LRE commuting

```{r}
m1_LREc <- gam(cbind(yes_batpass,no_batpass) ~
              s(Locality, bs = "re") +                       #bs = "re" is equivalent to adding Locality as random intercept in a mixed model (GAMM) ## smoothers should be 1 - keep tweaking the k value 
              s(jnight, by = Habitat, bs = "gp", k =90) +    
              Habitat +                                      
              s(temp_mean) +                               
              s(wind_mean) +                                 #same as for temp
              ti(temp_mean,wind_mean) ,                      #interaction between temp and wind
              data = LREc, method = "REML",                  #REML is not default, but is highly recommended by experts
              family = binomial, select=TRUE)                #may have to shift to quasibinomial if overdispersion, select = TRUE gives you automatic model selection

summary(m1_LREc) # 
# Now everything is significant , this model explains the data slightly better than the the feeding model as well. 
```
________________________________________________________________________________

Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 90) + Habitat + 
    s(temp_mean) + s(wind_mean) + ti(temp_mean, wind_mean)

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)         1.7274     0.2769   6.238 4.42e-10 ***
HabitatTurbinePad   0.9177     0.3003   3.056  0.00224 ** 
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                               edf Ref.df Chi.sq  p-value    
s(Locality)                 4.4688      6 15.480  0.00189 ** 
s(jnight):HabitatNatural    1.5270     87  9.367  0.00283 ** 
s(jnight):HabitatTurbinePad 4.1544     88 27.252 1.99e-05 *** 
s(temp_mean)                0.9738      9 32.126  < 2e-16 ***  # could be modeled as a linear effect
s(wind_mean)                3.7803      9 27.697  < 2e-16 ***
ti(temp_mean,wind_mean)     5.5758     16 38.951  < 2e-16 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.435   Deviance explained = 42.6%
-REML = 292.63  Scale est. = 1         n = 811

________________________________________________________________________________

```{r}
par(mfrow = c(2,2))
gam.check(m1_LREc)           #look at the plots, but also the output to check if k needs to be adjusted
gam.check(m1_LREc, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

overdispersion.m1_LREc <- sum( residuals(m1_LREc, "pearson")^2 ) / m1_LREc$df.residual
overdispersion.m1_LREc
# 0.7088693 ... is this too low? Yes... 

# From Katrine's Step3B GAM modelling script: 
#[1] 1.022701    #this value should ideally be 0.8-1.2, but it's not very bad!

```


# Basis dimension (k) checking results: 
The k-index should be approx. 1 or larger
Play play around with k....

```{r}

bmp(file.path(output_today, "checking k dimensions m1_LREc bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

windows()
plot(m1_LREc, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m1_LREc, scales ="free")         #function 'draw' is from package 'gratia'
draw(m1_LREc, scales ="fixed")

#dev.off() 
```


#Sketching the results - difference between Habitats
```{r}
levels(LREc$Habitat)
# "Natural"    "TurbinePad"

levels(LREc$Locality)
# [1] "Turbine10"           "Turbine11"           "Turbine14"          
# [5] "Turbine2"            "Turbine4"            "Turbine8"            "Turbine9"

pdata <- with(LREc,
              expand.grid(temp_mean=14,
                          wind_mean = 6,
                          Habitat = c("Natural", "TurbinePad"),
                          Locality = c("Turbine14"),
                          jnight = seq(min(jnight), max(jnight), 1))) #1 to keep the integer format

head(pdata)
tail(pdata)

fit <- data.frame(predict(m1_LREc, newdata=pdata, se.fit=TRUE, type = 'response'))
fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit))
pred <- cbind(pdata,fit)

head(pred)


bmp(file.path(output_today, "predictions m1_LREc bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

#windows()
plt2 <- ggplot(pred, aes(x=jnight, y = fit, group = factor(Habitat))) +
    
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) +
    geom_line() + facet_wrap(Locality ~ Habitat, scales ='free_y') +
    labs(x = "Julian night", y ="prop of hours per night with batpass>0")
plt2

dev.off()

```

# Now with SRE feeding 

```{r}
m1_SREf <- gam(cbind(yes_batpass,no_batpass) ~
              s(Locality, bs = "re") +                       #bs = "re" is equivalent to adding Locality as random intercept in a mixed model (GAMM)
              s(jnight, by = Habitat, bs = "gp", k =90) +    
              Habitat +                                      
              s(temp_mean) +                               
              s(wind_mean) +                                 #same as for temp
              ti(temp_mean,wind_mean) ,                      #interaction between temp and wind
              data = SREf, method = "REML",                  #REML is not default, but is highly recommended by experts
              family = binomial, select=TRUE)                #may have to shift to quasibinomial if overdispersion, select = TRUE gives you automatic model selection

summary(m1_SREf) # 
# The interaction between habitat and night is weak, as is the combined effect of wind and tempearature. Wind is still having a weird positive effect... 

```
________________________________________________________________________________

Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 90) + Habitat + 
    s(temp_mean) + s(wind_mean) + ti(temp_mean, wind_mean)

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)        -2.1860     0.3842  -5.690 1.27e-08 ***
HabitatTurbinePad  -1.0164     0.2370  -4.289 1.80e-05 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                                  edf Ref.df Chi.sq p-value    
s(Locality)                 5.224e+00      6 47.884 < 2e-16 ***
s(jnight):HabitatNatural    1.170e-04     87  0.000 0.99307    
s(jnight):HabitatTurbinePad 8.904e-05     87  0.000 0.83201    
s(temp_mean)                9.683e-01      9 34.763 < 2e-16 ***
s(wind_mean)                2.947e+00      9 12.172 0.00187 ** 
ti(temp_mean,wind_mean)     1.865e+00     16  4.475 0.05490 .  
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.309   Deviance explained = 33.7%
-REML = 257.62  Scale est. = 1         n = 811

________________________________________________________________________________

```{r}
par(mfrow = c(2,2))
gam.check(m1_SREf)           #look at the plots, but also the output to check if k needs to be adjusted
gam.check(m1_SREf, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

overdispersion.m1_SREf <- sum( residuals(m1_SREf, "pearson")^2 ) / m1_SREf$df.residual
overdispersion.m1_SREf
# 0.7270936 ... is this too low? # Yes, underdispersed, p values are a bit too high 
# The significant differences found are likely stronger than they seem 
# One solution may be to try a beta-distribution? googling 

# From Katrine's Step3B GAM modelling script: 
#[1] 1.022701    #this value should ideally be 0.8-1.2, but it's not very bad!

```


# Basis dimension (k) checking results: 
The k-index should be approx. 1 or larger
Play play around with k....

```{r}

bmp(file.path(output_today, "checking k dimensions m1_SREf bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

#windows()
plot(m1_SREf, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m1_SREf, scales ="free")         #function 'draw' is from package 'gratia'
draw(m1_SREf, scales ="fixed")

par(mfrow = c(1,2))
vis.gam(m1_SREf, view = c("temp_mean", "wind_mean"))    #Not a very elegant plot, you can find prettier solutions, but helps understand the interaction between tempp and wind
vis.gam(m1_SREf, view = c("wind_mean", "temp_mean")) 

dev.off() 
```


#Sketching the results - difference between Habitats
```{r}
levels(SREf$Habitat)
# "Natural"    "TurbinePad"

levels(SREf$Locality)
# [1] "Turbine10"           "Turbine11"           "Turbine14"          
# [5] "Turbine2"            "Turbine4"            "Turbine8"            "Turbine9"

# pdata <- with(SREf,
#               expand.grid(temp_mean=14,
#                           wind_mean = 6,
#                           Habitat = c("Natural", "TurbinePad"),
#                           Locality = c("Turbine14"),
#                           jnight = seq(min(jnight), max(jnight), 1))) #1 to keep the integer format
# 
# head(pdata)
# tail(pdata)

pdata <- with(SREf,
              expand.grid(temp_mean=14,
                          wind_mean = 6,
                          Habitat = c("Natural", "TurbinePad"),
                          Locality = c("Turbine14"),
                          jnight = seq(min(jnight), max(jnight), 1))) #1 to keep the integer format

head(pdata)
tail(pdata)

fit <- data.frame(predict(m1_SREf, newdata=pdata, se.fit=TRUE, type = 'response'))
fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit))
pred <- cbind(pdata,fit)

head(pred)


#bmp(file.path(output_today, "predictions m1_SREf bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

windows()
plt2 <- ggplot(pred, aes(x=jnight, y = fit, group = factor(Habitat))) +
    
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) +
    geom_line() + facet_wrap(Locality ~ Habitat, scales ='free_y') +
    labs(x = "Julian night", y ="prop of hours per night with batpass>0")
plt2

dev.off()





```

## Make predictions per Facility rather than Locality
## Make a new model that makes this possible. 
## Could use the function gam_predict() 




# Now with SRE commuting

```{r}
m1_SREc <- gam(cbind(yes_batpass,no_batpass) ~
              s(Locality, bs = "re") +                       #bs = "re" is equivalent to adding Locality as random intercept in a mixed model (GAMM)
              s(jnight, by = Habitat, bs = "gp", k =90) +    
              Habitat +                                      
              s(temp_mean) +                               
              s(wind_mean) +                                 #same as for temp
              ti(temp_mean,wind_mean) ,                      #interaction between temp and wind
              data = SREc, method = "REML",                  #REML is not default, but is highly recommended by experts
              family = binomial, select=TRUE)                #may have to shift to quasibinomial if overdispersion, select = TRUE gives you automatic model selection

summary(m1_SREc) # 
# 
```
________________________________________________________________________________


Family: binomial 
Link function: logit 

Formula:
cbind(yes_batpass, no_batpass) ~ s(Locality, bs = "re") + 
    s(jnight, by = Habitat, bs = "gp", k = 90) + Habitat + 
    s(temp_mean) + s(wind_mean) + ti(temp_mean, wind_mean)

Parametric coefficients:
                  Estimate Std. Error z value Pr(>|z|)    
(Intercept)        1.37150    0.27630   4.964 6.91e-07 ***
HabitatTurbinePad  0.06346    0.19587   0.324    0.746    
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Approximate significance of smooth terms:
                              edf Ref.df Chi.sq  p-value    
s(Locality)                 4.540      6 17.203 0.000811 ***
s(jnight):HabitatNatural    2.381     87 24.589 3.43e-06 ***
s(jnight):HabitatTurbinePad 3.608     88 43.538  < 2e-16 ***
s(temp_mean)                4.845      9 24.045 1.07e-05 ***
s(wind_mean)                3.045      9 44.741  < 2e-16 ***
ti(temp_mean,wind_mean)     1.239     16  2.866 0.052734 .  
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

R-sq.(adj) =  0.296   Deviance explained = 28.7%
-REML = 395.62  Scale est. = 1         n = 811

________________________________________________________________________________

```{r}
par(mfrow = c(2,2))
gam.check(m1_SREc)           #look at the plots, but also the output to check if k needs to be adjusted
gam.check(m1_SREc, rep=500)  #rep=500 gives you a polygon on the QQ plot, which the observed values should lie within

overdispersion.m1_SREc <- sum( residuals(m1_SREc, "pearson")^2 ) / m1_SREc$df.residual
overdispersion.m1_SREc
# 0.811134 ... good! 

# From Katrine's Step3B GAM modelling script: 
#[1] 1.022701    #this value should ideally be 0.8-1.2, but it's not very bad!

```


# Basis dimension (k) checking results: 
The k-index should be approx. 1 or larger
Play play around with k....

```{r}

bmp(file.path(output_today, "checking k dimensions m1_SREc bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

#windows()
plot(m1_SREc, pages=1,scheme=2,shade=TRUE,seWithMean = TRUE)

draw(m1_SREc, scales ="free")         #function 'draw' is from package 'gratia'
draw(m1_SREc, scales ="fixed")

dev.off() 
```


#Sketching the results - difference between Habitats
```{r}
levels(SREc$Habitat)
# "Natural"    "TurbinePad"

levels(SREc$Locality)
# [1] "Turbine10"           "Turbine11"           "Turbine14"          
# [5] "Turbine2"            "Turbine4"            "Turbine8"            "Turbine9"

pdata <- with(SREc,
              expand.grid(temp_mean=14,
                          wind_mean = 6,
                          Habitat = c("Natural", "TurbinePad"),
                          Locality = c("Turbine14"),
                          jnight = seq(min(jnight), max(jnight), 1))) #1 to keep the integer format

head(pdata)
tail(pdata)

fit <- data.frame(predict(m1_SREc, newdata=pdata, se.fit=TRUE, type = 'response'))
fit <- transform(fit, upper = fit + (2*se.fit), lower = fit-(2*se.fit))
pred <- cbind(pdata,fit)

head(pred)


bmp(file.path(output_today, "predictions m1_SREc bats_tot.jpeg"), width = 9, height = 6, units = "in", res = 350)

#windows()
plt2 <- ggplot(pred, aes(x=jnight, y = fit, group = factor(Habitat))) +
    
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'grey', alpha = 0.5) +
    geom_line() + facet_wrap(Locality ~ Habitat, scales ='free_y') +
    labs(x = "Julian night", y ="prop of hours per night with batpass>0")
plt2

dev.off()
```

### Notes from meeting with Katrine 29.11.2022

1. For the total bat data set, with y = binary nightly bat activity 
------------------------------------------------------------------------------------
a. Adjust the k value for the bats_tot model to find one that better suits the data
b. Illustrate the wind/temperature cutoffs. 
- At some point, it may be necessary to re-aggregate the weather data to Facility rather than across both facilities
------------------------------------------------------------------------------------

2. For the guild / behavior specific data sets 
------------------------------------------------------------------------------------
a. Find appropriate k values for each model type
b. Re run with y = bat pass count (bat pass sum, bat pass mean) rather than the binary response. 
c. Diagnose under dispersion where it exists 
d. Experiment with leaving out or including the interaction between wind and temperature
e. Make figures that show the overall activity at the different sites across time (perhaps include the mean or median as an added line with CI's) 
------------------------------------------------------------------------------------

3. For the bat / insect data
------------------------------------------------------------------------------------
a. Create continuous workflow from the bottom up and make it accessible on GitHub 
b. Have a dataset and a GitHub friendly R script that Katrine can use to start working with the data by 7. Dec 
------------------------------------------------------------------------------------

