---
title: "Marker bat manual acoustic analysis aggregated to night"
output: html_document
date: "2022-11-25"
---

# Input :  The cleaned manual acoustic analysis data from the Marker 2020 study 

# Outputs: 1. All bat activity aggregated to night per site 
           AND all bat activity aggregated to night per site - BINARY (1/0) 
           2. All guild and behavior bat activity aggregated to night per site
           AND all bat activity aggregated to night per site - BINARY (1/0) 
           

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(tidy = "styler")
options(knitr.table.format = "html")

# set project working directory according to the user system info 
# Otherwise, all the data (inputs and outputs, including figures) can be stored on a shared OneDrive folder

user <- Sys.info()['effective_user'] 
user
# this should print your nmbu user name - "apmc" in my case. 

wd <- getwd()
wd
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/Documents/1. PhD_Main/GitHub_link/MarkerBatAcousticAnalyses/MarkerBatAcousticAnalyses2022"


##########################################################
#### Work environment set up ####
##########################################################

library(knitr)
library(data.table)
library(tidyverse)
library(beepr)
library(lubridate)
library(purrr)
#renv::install("rstudio/renv")
library(renv)
library(stringr)
library(janitor)
library(anytime)
library(kableExtra)
library(papeR)
library(skimr)
library(vtable)

##########################################################
#### Import data  ####
##########################################################

# All input data can be found on a shared OneDrive folder - we can both share the same input folder but we should have different Output folders. 

# for Katrine 
#input <- ""

# for Reed 
input <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Input"

dataset1 <- "MarkerManual_cleaned_manualid_guild_sitenames.csv"
dataset2 <- "Marker_night_inventory_0704.2022_edit.csv"
dataset3 <- "nightlyaggregate_wind_temp.csv"

path1 <- str_c(input, "/", dataset1)
path2 <- str_c(input, "/", dataset2)
path3 <- str_c(input, "/", dataset3)

bats1. <- read_csv(path1) # 19438 bat passes with 22 vars
active.nights <- read.csv(path2, sep = ";") # # 951 obs of 3 variables
weather.nights <- read_csv(path3) # 93 obs of 8 vars


# for Reed
output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Reed/Outputs"

# for Katrine 
#output <- ""

## 
 file.name <- "Marker aggregated to night_all bats"
# 
 todays_date <- Sys.Date()
# 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
 dir.name
# 
 dir.create(dir.name)

output_today <- dir.name

```

# Make some pretty tables with the unaggregated dataset 
```{r}
# make factor levels arranged in orders that are more meaningful and insert a taxa column with more specific manual id names. 

names(bats1.)
cols <- c("manual.id" ,  "behavior"  ,  "guild"     ,  
          "Site"     ,   "Habitat"   ,  "Locality"  ,  "Facility")
bats1.[cols] <- lapply(bats1.[cols], factor)
bats1.$taxa <- bats1.$manual.id
levels(bats1.$taxa)
# "BABA" "EPNI" "LR1"  "LR2"  "MR1"  "NYNO" "NoID" "PAUR" "PINA" "PIPY" "SR1"  "VEMU"
levels(bats1.$taxa) <- list("Barbastella barbastellus" = "BABA",
                           "Plecotus auritus" = "PAUR",
                           "Short range group 1" = "SR1",
                           "Eptesicus nilssonii" = "EPNI",
                           "Nyctalus noctula" = "NYNO",
                           "Vespertilio murinus" = "VEMU",
                           "Long range group 1" = "LR1",
                           "Long range group 2" = "LR2",
                           "Medium range group1" = "MR1",
                           "Pipistrellus nathusii" = "PINA",
                           "Pipistrellus pygmaeus" = "PIPY",
                            "Unknown bat" = "NoID") 
summary(bats1.$taxa)

levels(bats1.$guild) <- list("SRE" = "SRE", "MRE" = "MRE", "LRE" = "LRE", "NoID" = "NoID")
summary(bats1.$guild)

levels(bats1.$Locality)
levels(bats1.$Locality) <-list("Turbine2" = "Turbine2" , "Turbine4" = "Turbine4",  "Turbine8" = "Turbine8", "MeteorologicalTower" = "MeteorologicalTower", "Turbine9" = "Turbine9", "Turbine10" = "Turbine10", "Turbine11" = "Turbine11", "Turbine14" = "Turbine14") 
summary(bats1.$Locality)

# Can play with these and make them much nicer... 
kbl(summarize(bats1., type = "factor", variables = "manual.id"))  %>%  kable_styling()
kbl(summarize(bats1., type = "factor", variables = "taxa"))  %>%  kable_styling()
kbl(summarize(bats1., type = "factor", variables = "guild"))  %>%  kable_styling()
kbl(summarize(bats1., type = "factor", variables = "Site"))  %>%  kable_styling()
kbl(summarize(bats1., type = "factor", variables = "Habitat"))  %>%  kable_styling()
kbl(summarize(bats1., type = "factor", variables = "Facility"))  %>%  kable_styling()

names(bats1.)
bats2. <- bats1. %>% select(behavior, guild, Site, Habitat, Locality, Facility, taxa) 

st(bats2.)

# will find a way to automatically print this table later... 
# dev.print(device = .png, 
#           filename = file.path(output_today, "summarytable.png"),
#           width = 480, 
#           height = 300)
```


1. all bat activity aggregated to night per site
- Insert zeros
- merge with night aggregated weather data 
```{r}
bats3. <- bats1. %>% mutate(date = as.Date(DATE),
                           night = as.Date(DATE.12))

bat_sitemap <- bats1. %>% select(Site, Habitat, Facility) %>% distinct() 

bats_nights_sum <- bats3. %>% group_by(Site, night) %>% dplyr::summarize(batpass = sum(n())) 
bats_nights_sum1 <- left_join(bats_nights_sum, bat_sitemap)
summary(bats_nights_sum1)
summary(bats_nights_sum1$Site)

#write.csv(bats_nights_sum1, "C:/Users/April/OneDrive - Norwegian University of Life Sciences/Documents/1. PhD_Main/GitHub/Marker_Acoustic_Analyses/From Manual output to nightly aggregated/Outputs/Outputs_nightaggregated/bat pass aggregated by night wout zeronights.csv")

################################################
# Insert zero activity nights
################################################
head(active.nights)

active.nights$Site <- gsub("C", "N", active.nights$Site)
active.nights$Site <- gsub("MetA", "Met45", active.nights$Site)
active.nights$Site <- gsub("MetB", "Met95", active.nights$Site)
active.nights$Site <- as.factor(active.nights$Site)
active.nights$night <- as.IDate(active.nights$night, '%d.%m.%Y')
active.nights$night <- as.Date(active.nights$night)
summary(active.nights)

dat <- bats_nights_sum1
summary(dat)
summary(dat$Site)
#---- ADDING MISSING DAYS 

# adjust dataset to include nights when detectors were active but no data was collected. 

head(active.nights)
summary(active.nights$Site)

nights <- unique(active.nights$night) # 91 unique nights of survey 

temp<-expand.grid(Site = unique(dat$Site),  night = nights) #----expanded data
temp$night <- as.character(temp$night)
temp$night <-  as.Date(temp$night)

active.nights$Site <- as.factor(active.nights$Site)
active.nights$night <-  as.Date(active.nights$night)

str(temp)
str(active.nights)

temp2<-merge(temp, active.nights) # 811 obs of 3 variables 

summary(temp2$Site) # number of monitoring nights per site 
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#    70    70    59    44    60    48    60    60    60    60    60    60    60    60    60    60 

temp3<-merge(temp2,dat,all.x=TRUE) # 811 obs of 6 vars 
summary(temp3$Site)# number of nights per guild per site plot 
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#    70    70    59    44    60    48    60    60    60    60    60    60    60    60    60    60 

## Compare to the original dataset
summary(dat$Site)
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#    55    17    54    41    49    46    50    55    51    47    52    49    55    49    54    54 

summary(temp3) # 173 zero nights introduced 


# reformat factor columns and then replace NA bat passes with zeros 
nbats1 <- temp3 

#Replace NAs with zeros 
nbats1[is.na(nbats1)] <- 0 
summary(nbats1)

# batpasses converted to zeros but NAs pop up in the factor columns
# fix Habitat and Facility 

nbats2 <- nbats1 %>% select(-c(active.night, Habitat, Facility)) 
summary(nbats2)  

nbats3 <- left_join(nbats2, bat_sitemap) 
summary(nbats3)

#write.csv(nbats3, file.path(output_today,"nightlyaggregatedMarker2020_totalbats_zeroinserted.csv")) 

```

# Aggregate the hourly weather datasets to nightly 
```{r}
head(weather.nights)

nbats4 <- left_join(nbats3, weather.nights, by = "night")

summary(nbats4)

# right now I do not have UTM or day length included but I can add those here later. 

```


# Make it binary! 
```{r}
bats5 <- nbats4 %>% rename(Batpass = batpass)
bats5$batpass01 <- bats5$Batpass
names(bats5)
 # [1] "Site"        "night"       "batpass_sum" "Habitat"     "Facility"    "avg_wind"    "min_wind"    "max_wind"   
 # [9] "avg_temp"    "min_temp"    "max_temp"    "batpass01"  
bats5$batpass01[bats5$batpass01 >1] <- 1
summary(bats5)
head(bats5)
tail(bats5)

#write.csv(bats5, file.path(
#  output_today,"nightlyaggregatedMarker2020_totalbats_zeroinserted_weather_binary.csv")) 

# last rounds of housekeeping

bats5$wind <- bats5$avg_wind
bats5$temp <- bats5$avg_temp
bats5$jnight <- yday(bats5$night)

# Aggregation table 
# I am keeping the night (as.Date) variable for now because it is easier to recreate jnight from night than the other way around. 

agg_tbl2 <- bats5 %>% group_by(night, Site) %>% 
  dplyr::summarise(across(c(Batpass, batpass01, wind, temp), list(sum = sum, mean = mean, min = min, max=max, length = length)))

dim(agg_tbl2)
# 951  22
head(agg_tbl2)
tail(agg_tbl2)

#Remove variables that will not be needed
names(agg_tbl2)

#  [1] "jnight"           "Site"             "Batpass_sum"      "Batpass_mean"     "Batpass_min"      "Batpass_max"     
#  [7] "Batpass_length"   "batpass01_sum"    "batpass01_mean"   "batpass01_min"    "batpass01_max"    "batpass01_length"
# [13] "wind_sum"         "wind_mean"        "wind_min"         "wind_max"         "wind_length"      "temp_sum"        
# [19] "temp_mean"        "temp_min"         "temp_max"         "temp_length"     


df2C <- agg_tbl2 %>% select(-c(Batpass_min, Batpass_length, batpass01_mean, batpass01_min, batpass01_max,
                                wind_sum, wind_length, temp_sum , temp_length ))

dim(df2C) 
# 951  13
head(df2C)
tail(df2C)


#Make a proportion variable

df2C$batpass_prop <- df2C$batpass01_sum/df2C$batpass01_length

head(df2C)
tail(df2C)
dim(df2C)
summary(df2C) # is also only ones or zeros... why apply this to the binary variable instead of the bat pass sum? 
sum(df2C$Batpass_sum)
# 19438 we have the correct number of bat passes so this looks good! 

#write.csv(df2C, file.path(output_today, "totalbatpass_summarytable_withbinary_batpass_night_aggregated.csv")) 

# Add back the other site data 

df3C <- left_join(df2C, bat_sitemap)

#write.csv(df3C, file.path(output_today,"totalbatpass_summarytable_withbinary_batpass_night_aggregated_site data.csv"))

```


### Do some quality control plots 

```{r}
names(df3C)

#batpass_prop	versus	jnight
# save plot to file 
bmp(file.path(output_today,"Habitat_batpass_prop_night.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("night") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df3C, 
                    aes(x = night, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df3C, 
                     aes(x = night, 
                         y = batpass_prop))

p <- p + facet_grid( ~ Habitat)


print(p)
dev.off()

 

#batpass_prop	versus	temp_max			
# save plot to file 
bmp(file.path(output_today, "Habitat_batpass_prop_temp_max.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("temp_max") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df3C, 
                    aes(x = temp_max, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df3C, 
                     aes(x = temp_max, 
                         y = batpass_prop))

p <- p + facet_grid( ~ Habitat)


print(p)
dev.off()

#batpass_prop	versus	temp_mean			
# save plot to file 
bmp(file.path(output_today, "Habitat_batpass_prop_temp_mean.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("temp_mean") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df3C, 
                    aes(x = temp_mean, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df3C, 
                     aes(x = temp_mean, 
                         y = batpass_prop))

p <- p + facet_grid( ~ Habitat)


print(p)
dev.off()

#batpass_prop	versus	wind_max			
# save plot to file 
bmp(file.path(output_today, "Habitat_batpass_prop_wind_max.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("wind_max") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df3C, 
                    aes(x = wind_max, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df3C, 
                     aes(x = wind_max, 
                         y = batpass_prop))

p <- p + facet_grid( ~ Habitat)


print(p)
dev.off()

#batpass_prop	versus	wind_mean			
# save plot to file 
bmp(file.path(output_today, "Habitat_batpass_prop_wind_mean.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("wind_mean") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df3C, 
                    aes(x = wind_mean, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df3C, 
                     aes(x = wind_mean, 
                         y = batpass_prop))

p <- p + facet_grid( ~ Habitat)


print(p)
dev.off()

```

###############################################################################

I am going to recycle some names and data objects from above. This is not best practice and I should probably separate this into two separate markdowns later or at least change the names of the objects below but for now I am just going to run with it. 


# 2. Aggregated to night by guild and behavior 

```{r}
# bats3. <- bats1. %>% mutate(date = as.Date(DATE),
#                            night = as.Date(DATE.12))
# 
# bat_sitemap <- bats1. %>% select(Site, Habitat, Facility) %>% distinct() 

bats_gb_nights_sum <- bats3. %>% group_by(Site, night, guild, behavior) %>% dplyr::summarize(batpass = sum(n()))
summary(bats_gb_nights_sum)
#bats_gb_nights_sum1 <- left_join(bats_gb_nights_sum, bat_sitemap) # can skip this, the site data will be added back at the end
bats_gb_nights_sum1 <- bats_gb_nights_sum
summary(bats_gb_nights_sum1)
summary(bats_gb_nights_sum1$Site)
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#    68    20   140   132   142   167   181   138   158   134   117   114   185   162   174   139 

#write.csv(bats_gb_nights_sum1, file.path(output_today, "bat pass aggregated by night wout zeronights_behavior and guild.csv")) 

################################################
# Insert zero activity nights
################################################

# same active.nights objects as above should work fine! 
summary(active.nights)

dat <- bats_gb_nights_sum1
summary(dat)
names(dat)
# "Site"     "night"    "guild"    "behavior" "batpass" 

#---- ADDING MISSING DAYS 

# adjust dataset to include nights when detectors were active but no data was collected. 

head(active.nights)
summary(active.nights$Site)

nights <- unique(active.nights$night) # 91 unique nights of survey 

temp<-expand.grid(Site = unique(dat$Site), 
                  guild = unique(dat$guild), 
                  behavior = unique(dat$behavior),  
                  night = nights) #----expanded data

temp$night <- as.character(temp$night)
temp$night <-  as.Date(temp$night)

active.nights$Site <- as.factor(active.nights$Site)
active.nights$night <-  as.Date(active.nights$night)

str(temp)
str(active.nights)

temp2<-merge(temp, active.nights) # 811 obs of 3 variables 

summary(temp2$Site) # number of monitoring nights per site 
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#   840   840   708   528   720   576   720   720   720   720   720   720   720   720   720   720 

temp3<-merge(temp2,dat,all.x=TRUE) # 811 obs of 6 vars 
summary(temp3$Site)# number of nights per guild per site plot 
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#   840   840   708   528   720   576   720   720   720   720   720   720   720   720   720   720 

## Compare to the original dataset
summary(dat$Site)
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#    68    20   140   132   142   167   181   138   158   134   117   114   185   162   174   139 

summary(temp3) # 9241 zero nights/behavior/guilds introduced 


# reformat factor columns and then replace NA bat passes with zeros 
nbats1 <- temp3 

#Replace NAs with zeros 
nbats1[is.na(nbats1)] <- 0 
summary(nbats1)

sum(nbats1$batpass) 
# 19438 # good! 

# batpasses converted to zeros but NAs pop up in the factor columns
# fix Habitat and Facility 

nbats2 <- nbats1 %>% select(-c(active.night)) 
summary(nbats2)  

nbats3 <- left_join(nbats2, bat_sitemap) 
summary(nbats3) # 11412 obs of 13 variables 

#write.csv(nbats3, file.path(output_today, "nightlyaggregatedMarker2020_totalbats_zeroinserted_behavior and guild.csv")) 

```

# Aggregate the hourly weather datasets to nightly 
```{r}
head(weather.nights)

nbats4 <- left_join(nbats3, weather.nights, by = "night")

summary(nbats4)

# right now I do not have UTM or day length included but I can add those here later. 

```


2. Make it binary! 
```{r}
bats5 <- nbats4 %>% rename(Batpass = batpass)
bats5$batpass01 <- bats5$Batpass
names(bats5)
 # [1] "Site"        "night"       "batpass_sum" "Habitat"     "Facility"    "avg_wind"    "min_wind"    "max_wind"   
 # [9] "avg_temp"    "min_temp"    "max_temp"    "batpass01"  
bats5$batpass01[bats5$batpass01 >1] <- 1
summary(bats5)
head(bats5)
tail(bats5)


#write.csv(bats5, file.path(output_today, "nightlyaggregated_zeros_binary_behavior and guild.csv")) 

# last rounds of housekeeping

bats5$wind <- bats5$avg_wind
bats5$temp <- bats5$avg_temp
bats5$jnight <- yday(bats5$night)

# Aggregation table 
# I am keeping night rather than julian night because it is easy to recreate julian night later but converting julian night back to night is a bit trickier. 
agg_tbl2 <- bats5 %>% group_by(night, Site, guild, behavior) %>% 
  dplyr::summarise(across(c(Batpass, batpass01, wind, temp), list(sum = sum, mean = mean, min = min, max=max, length = length)))

dim(agg_tbl2)
# 11412    24
head(agg_tbl2)
tail(agg_tbl2)

#Remove variables that will not be needed
names(agg_tbl2)

#  [1] "jnight"           "Site"             "guild"            "behavior"         "Batpass_sum"      "Batpass_mean"    
#  [7] "Batpass_min"      "Batpass_max"      "Batpass_length"   "batpass01_sum"    "batpass01_mean"   "batpass01_min"   
# [13] "batpass01_max"    "batpass01_length" "wind_sum"         "wind_mean"        "wind_min"         "wind_max"        
# [19] "wind_length"      "temp_sum"         "temp_mean"        "temp_min"         "temp_max"         "temp_length"   


df2C <- agg_tbl2 %>% select(-c(Batpass_min, Batpass_length, batpass01_mean, batpass01_min, batpass01_max,
                                wind_sum, wind_length, temp_sum , temp_length ))

dim(df2C) 
# 11412    15
head(df2C)
tail(df2C)


#Make a proportion variable

df2C$batpass_prop <- df2C$batpass01_sum/df2C$batpass01_length

head(df2C)
tail(df2C)
dim(df2C)
summary(df2C) # is also only ones or zeros... why apply this to the binary variable instead of the bat pass sum? 
sum(df2C$Batpass_sum)
# 19438 we have the correct number of bat passes so this looks good! 

#write.csv(df2C, file.path(output_today, "totalbatpass_summarytable_withbinary_batpass_night_aggregated_behavior and guild.csv")) 

# Add back the other site data 

df3C <- left_join(df2C, bat_sitemap)
head(df3C)

#write.csv(df3C, file.path(output_today,"totalbatpass_summarytable_site.csv")) 

```



```{r}
# summary of survey effort
names(df3C)
#  [1] "jnight"           "Site"             "guild"            "behavior"         "Batpass_sum"      "Batpass_mean"     "Batpass_max"     
#  [8] "batpass01_sum"    "batpass01_length" "wind_mean"        "wind_min"         "wind_max"         "temp_mean"        "temp_min"        
# [15] "temp_max"         "batpass_prop"     "Habitat"          "Facility"    
summary(df3C)


bmp(file.path(output_today,"Survey effort per site.tif"), width = 9, height = 6, units = "in", res = 350)
# eventually convert julian night to a proper date 
p <- ggplot(data=df3C[df3C$Batpass_sum>0,]) +
  geom_point(data=df3C, aes(x=night,y=guild),shape=19,color="grey") +
  geom_point(aes(x=night,y=guild),shape=19,color="black") +
  facet_wrap(~Site)
print(p)
dev.off()

p
```

```{r}
# Remove Met tower , MRE bats, and social calls. 

df1 <- df3C %>%  
  dplyr::filter(Habitat != "MeteorologicalTower") %>% 
  dplyr::filter(guild != "MRE") %>% 
  dplyr::filter(behavior != "Social") %>% 
  droplevels() 

summary(df1) # 4866 obs of 18 vars

#write.csv(df3C, file.path(output_today, "totalbatpass_summarytable_trimmed MRE social and met tower.csv")) 
### copied and renamed as "guild_behavior_batpass_summarytable_trimmed MRE social and met tower"

```

## New summary plots 

```{r}
names(df1)

#  [1] "night"            "Site"             "guild"            "behavior"         "Batpass_sum"      "Batpass_mean"    
#  [7] "Batpass_max"      "batpass01_sum"    "batpass01_length" "wind_mean"        "wind_min"         "wind_max"        
# [13] "temp1_mean"        "temp1_min"         "temp1_max"         "batpass_prop"     "Habitat"          "Facility"   

#batpass_prop	versus	night
# save plot to file 
bmp(file.path(output_today, "Habitat_batpass_prop_night.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("night") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df1, 
                    aes(x = night, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df1, 
                     aes(x = night, 
                         y = batpass_prop))

p <- p + facet_grid( behavior ~ Habitat)


print(p)
dev.off()

 

#batpass_prop	versus	temp1_max			
# save plot to file 
bmp(file.path(output_today,"Habitat_batpass_prop_temp1_max.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("temp_max") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df1, 
                    aes(x = temp_max, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df1, 
                     aes(x = temp_max, 
                         y = batpass_prop))

p <- p + facet_grid( behavior ~ Habitat)


print(p)
dev.off()

#batpass_prop	versus	temp1_mean			
# save plot to file 
bmp(file.path(output_today,"Habitat_batpass_prop_temp1_mean.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("temp_mean") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df1, 
                    aes(x = temp_mean, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df1, 
                     aes(x = temp_mean, 
                         y = batpass_prop))

p <- p + facet_grid(behavior ~ Habitat)


print(p)
dev.off()

#batpass_prop	versus	wind_max			
# save plot to file 
bmp(file.path(output_today,"Habitat_batpass_prop_wind_max.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("wind_max") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df1, 
                    aes(x = wind_max, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df1, 
                     aes(x = wind_max, 
                         y = batpass_prop))

p <- p + facet_grid(behavior ~ Habitat)


print(p)
dev.off()

#batpass_prop	versus	wind_mean			
# save plot to file 
bmp(file.path(output_today,"Habitat_batpass_prop_wind_mean.tif"), width = 9, height = 6, units = "in", res = 350)

p <- ggplot()
p <- p + xlab("wind_mean") + ylab("batpassTOTAL_prop")
p <- p + theme(text = element_text(size = 15))
p <- p + geom_point(data = df1, 
                    aes(x = wind_mean, 
                        y = batpass_prop),
                    size = 1)
p <- p + geom_smooth(data = df1, 
                     aes(x = wind_mean, 
                         y = batpass_prop))

p <- p + facet_grid(behavior ~ Habitat)


print(p)
dev.off()
```


